Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2020-11-19 paper/main.tex
one ref left

Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$. However, in other case uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$. Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain. For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
%
% In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$. Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
Interestingly, uBorda is computed in polynomial time.

@@

\subsection{CEGO}

Combinatorial Efficient Global Optimization
(CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
fitness function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

@@

Moreover, we consider real LOP datasets from the LOPLIB
repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
particular, we consider the following types: IO, Random instances of type A and
Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
250\}$. These are harder problems than the synthetic instances, not only
because of the larger search space, but also because of the different instance
types.

@@

\begin{figure}[htb]
  \centering%
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d100.01} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d50.01} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC05} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_rec05}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{kra32} (QAP)\\[-0.5ex]
\includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
\end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better). Whitespace denotes missing data (runs that took longer than 6 hours).\label{fig:heatmaps}}
\end{figure}

@@

To better understand the behavior of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an
example, each line in Fig.~\ref{fig:lop_synth} shows the mean fitness of the
solutions evaluated by either UMM or CEGO on one synthetic LOP instance. The
shaded area shows one standard deviation around the mean. Ideally, each new
evaluation will monotonically improve in fitness, since each solution evaluated
helps to refine the model and leads to a better solution being evaluated in the
next step.  However, it is often the case that new solutions are often worse
than their predecessors. Nevertheless, in this example, there is a clear
overall improvement as more evaluations are performed. The minimum values
achieved in the plot up to a particular number of function evaluations gives an
estimation of the mean fitness of the best-so-far solution up to that point in
the run. The algorithms do not return the final solution shown but rather the
best one (minimum fitness) of each run.

@@

When we compare the behavior of CEGO and
UMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}), we observe a clear
pattern independently of the uniformity of the instances. In particular, we see
a linear convergence to the optimal value for UMM.  This is due to the linear
decrease in the parameter $\theta$ (line~\ref{line:theta} in
Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
internal probability model of the algorithm over the $\FEmax$ functions
evaluations. Overall, the mean fitness returned by UMM after $\FEmax$ is better one returned by  CEGO, regardless the uniformity of the instances.

@@

On the other hand, % \EKHINE{cambio el orden}
CEGO starts with a random sampling of 10 solutions, with are quite poor as
expected, followed by building and updating the Kriging model based new
evaluations, which leads to an impressive improvement in less than 50
evaluations. However, progress is halted at this point and the model seems to
have trouble generating better solutions. Around evaluation 200, the search
becomes fully random. Lower uniformity values, e.g., 0.5, show worsening
evaluations despite the increase in solutions used to build the model. In
summary, the Kriging model clearly helps CEGO to quickly identify good
solutions in very few evaluations. However, for some reason, CEGO is not able
to further improve those solutions given more data. Given what we know about
Kriging~\citep{EriPeaGar2019scalable}, it appears that the surrogate model cannot model the local step
necessary to reach the optimal solution and switches to an exploration strategy
that ends up being completely random.

@@

%: LOP reales: he puesto 3 de diferentes tamanos. Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
\begin{figure}[tbp]
  \centering%
  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandB_N-p50-01}\\
  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d100_01}\\
  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d150_01}\\
%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
  \end{figure}

%Results are quite different on the 
%The fast initial convergence of CEGO for synthetic LOP instances does not seem
%to translate to the
%instances available in LOPLIB~\citep{}. We show in
%Fig.~\ref{fig:loplib} three examples with different permutation size, however,
%the results are consistent for other LOPLIB instances (complete results
%available as supplementary material \supplement). Due to extremely long
%runtimes of CEGO and constraints in our computing system, we set a maximum
%wall-clock time of 3 days for each run of CEGO. On these instances, the
%behaviour of CEGO and UMM is much more similar (at least up to the termination
%point of CEGO). It appears that our synthetic LOP instances have a fitness
%landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
%instances do not present such landscape. Comparing CEGO and UMM results, the
%differences get smaller, in favour of UMM, with larger instance size. Our
%conjecture is that building an accurate Bayesian model and searching for good
%solutions on it becomes harder for larger fitness landscapes.
%\EKHINE{Parece que el paper va del CEGO en vez del UMM :p. Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
%\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
%\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}

@@

In the following paragraphs we analyze the result over real-world instances. %
Figure~\ref{fig:loplib} shows the results on the instances taken from LOPLIB.
  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
The type of instances from the synthetic LOP instances, and, thus, so does the
landscape. However, UMM shows a very similar behavior: there is a (close to)
linear decrease of the fitness of new solutions over each new evaluation. This
suggests that UMM is able to model the fitness landscape. Regarding CEGO, we
observe a similar behavior for the first 50-100 evaluations. At this point,
CEGO reaches the time limit of 3 days, so no observations can be made for
longer budgets.  Overall, the improvement from the first (random) to the last
solutions is 2-3 times larger in UMM than CEGO. %
%\MANUEL{But this is a bit misleading because UMM runs for 4 times more
%  evaluations, perhaps we should also measure up to the evaluation that CEGO
%  reaches so we report both values?}

@@

In this paper, we have introduced UMM, a population-based probabilistic
algorithm based on an unbalanced Mallows model. The algorithm is designed for
black-box combinatorial problems on permutation landscapes and when the budget
of fitness evaluations is severely limited (here, up to $400$ evaluations).
UMM \ekhine{}{has shown to perform statistically better or equal than CEGO, the
  state of the art algorithm for black-box expensive optimization, in the same instances as in the original paper of
  CEGO and some new instances. UMM} is specially well-suited for budget-limited
scenarios that are still time-sensitive, e.g., when the overhead incurred by
the optimizer should be no more than a few minutes added to each fitness
evaluation.

%\MANUEL{UMM is able to model the LOP landscape
%  at least as well as CEGO as shown by the plots where UMM shows a linear
%  decrease in fitness.}  
%  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
%  solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q
%  es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos y
%  luego que addemás es más rápido} 
%  \MANUEL{Pero UMM es peor en rec13 y nug12, y
%  en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más
%  rápido. No podemos saber qué pasaría si CEGO llegara a 400. Podríamos medir
%  lo que pasa en el momento en que matamos a CEGO y dar también ese valor
%  entonces la diferencia en velocidad no importaría. } 
The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
CEGO  it sharply
increases with $n$, requiring 
sometimes more than hour per fitness evaluation as shown in the experimental section. 
Hence, UMM is a computationally
feasible alternative for relatively large problem sizes.

Although we were inspired by previous work on ACO for expensive black-box
combinatorial problems~\citep{PerLopStu2015si}, UMM is, to the best of our
knowledge, the first EDA specifically designed for such problems that is able
to match (and sometimes even improve) the results of a Bayesian
optimizer.
The results presented here show that, despite the intrinsic difficulty of the
expensive black-box combinatorial scenario, there are still significant
advances yet to be made. Thus, we hope that this paper will motivate further
research. % on such problems.
%\MANUEL{I'm thinking about finishing here and commenting out the following paragraph}
A somewhat surprising result was that the behavior of UMM (and the same for
CEGO) on the PFSP is close to random. This result may be due to the distance
metric used, but further analysis is needed to corroborate this
conjecture. Here, we have focused on Kendall's-$\tau$ distance, however, it is
possible to extend UMM to other distance metrics, which will allow us to
dynamically select among various distance metrics for an unknown black-box
% \ekhine{}{estoy totalmente en contra de esto :) }
permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
always obvious, thus a more detailed analysis would be needed to provide either
generally good static values or an online adaptation approach. Finally, we have
used here three very different combinatorial problems (LOP, PFSP and QAP) as
black-box benchmarks. However, an even more diverse range of problems would be
needed to understand the behavior of UMM on real-world black-box combinatorial
landscapes.

@@

%% MANUEL: Blind, so no acknowledgments yet.
\begin{smaller}
  \paragraph*{Reproducibility.} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
  \vspace*{-1em}
\paragraph*{Acknowledgements.}
% M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
%
\manuel{}{: We should also thank the COST ACTION: Ask Carlos?} \ekhine{}{Sure!, esto no habra q ponerlo aun por ser blind, no?}
Thanks to Hao Wang (Leiden University) for pointing us to the arguments of
\citet{EriPeaGar2019scalable} and to Thomas Stützle (Université libre de Bruxelles) for suggesting works
discussing differences between QAPLIB instances.
Thanks to Ahmed Boujaada for the experimental support. 
\end{smaller}

\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
\bibliographystyle{splncs04nat}
\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}

>>>>

Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$. Otherwise, uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$. Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
The choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
%% MANUEL: To save space.
%For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
%
% In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$. Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
Finally, uBorda is computed in polynomial time.

@@

\subsection{Combinatorial Efficient Global Optimization (CEGO)}

CEGO~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
fitness function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

@@

Moreover, we consider real LOP datasets from the LOLIB
repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
particular, we consider the following types: IO, Random instances of type A and
Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
250\}$. These are harder problems than the synthetic instances, not only
because of the larger search space, but also because of the different instance
types.

@@

\begin{figure}[tb]
  \centering%
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d100.01} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d50.01} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC05} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_rec05}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{kra32} (QAP)\\[-0.5ex]
\includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
\end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better). Whitespace denotes missing data (runs that took longer than 6 hours).\label{fig:heatmaps}}
\end{figure}

@@

To better understand the behavior of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an
example, each line in Fig.~\ref{fig:lop_synth} shows the mean fitness of the
solutions evaluated by either UMM or CEGO on one synthetic LOP instance. The
shaded area shows one standard deviation around the mean. Ideally, each new
evaluation will monotonically improve in fitness, since each solution evaluated
helps to refine the model and leads to a better solution being evaluated in the
next step.  However, it is possible that new solutions are worse
than their predecessors. %Nevertheless, in this example, there is a clear
% overall improvement as more evaluations are performed.
The minimum values
achieved in the plot up to a particular number of function evaluations gives an
estimation of the mean fitness of the best-so-far solution up to that point in
the run. The algorithms do not return the final solution shown but rather the
best one (minimum fitness) of each run.

@@

When we compare the behavior of CEGO and
UMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}), we observe a clear
pattern independently of the uniformity of the instances. In particular, we see
a linear convergence to the optimal value for UMM.  This is due to the linear
decrease in the parameter $\theta$ (line~\ref{line:theta} in
Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
internal probability model of the UMM over the $\FEmax$ functions
evaluations. Overall, the mean fitness returned by UMM at $\FEmax=400$ is better than the one returned by  CEGO, regardless the uniformity of the instances.

@@

On the other hand, % \EKHINE{cambio el orden}
CEGO starts with 10 random initial solutions, % with are quite poor as
% expected,
followed by building and updating the Kriging model based on new
evaluations, which leads to an impressive improvement in less than 50
evaluations. However, progress is halted at this point and CEGO seems to
have trouble finding better solutions. Around evaluation 200, the search
becomes fully random. Lower uniformity values, e.g., 0.5, show worsening
evaluations despite the increase in solutions used to build the model. In
summary, the Kriging model clearly helps CEGO to quickly identify good
solutions in very few evaluations. However, CEGO is not able
to further improve those solutions given more data. Given what we know about
Kriging~\citep{EriPeaGar2019scalable}, it appears that the surrogate model cannot model the local step
necessary to reach the optimal solution and switches to an exploration strategy
that ends up being completely random.

@@

%: LOP reales: he puesto 3 de diferentes tamanos. Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
\begin{figure}[!tbp]
  \centering%
  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandB_N-p50-01}\\
  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d100_01}\\
  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d150_01}\\
%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOLIB.\label{fig:lolib}}
  \end{figure}

%Results are quite different on the 
%The fast initial convergence of CEGO for synthetic LOP instances does not seem
%to translate to the
%instances available in LOLIB~\citep{}. We show in
%Fig.~\ref{fig:lolib} three examples with different permutation size, however,
%the results are consistent for other LOLIB instances (complete results
%available as supplementary material \supplement). Due to extremely long
%runtimes of CEGO and constraints in our computing system, we set a maximum
%wall-clock time of 3 days for each run of CEGO. On these instances, the
%behaviour of CEGO and UMM is much more similar (at least up to the termination
%point of CEGO). It appears that our synthetic LOP instances have a fitness
%landscape that is very amenable to CEGO's Kriging model, whereas the LOLIB
%instances do not present such landscape. Comparing CEGO and UMM results, the
%differences get smaller, in favour of UMM, with larger instance size. Our
%conjecture is that building an accurate Bayesian model and searching for good
%solutions on it becomes harder for larger fitness landscapes.
%\EKHINE{Parece que el paper va del CEGO en vez del UMM :p. Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
%\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
%\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}

@@

In the following paragraphs we analyze the results over real-world instances. %
Figure~\ref{fig:lolib} shows the results on the instances taken from LOLIB.
  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
The type of instances differs greatly from the synthetic LOP instances, and,
thus, so does the landscape. However, UMM shows a very similar behavior: there
is a (close to) linear decrease of the fitness of new solutions over each new
evaluation. This suggests that UMM is able to model the fitness
landscape. Regarding CEGO, we observe a similar behavior to UMM for the first
50-100 evaluations, however, in several instances CEGO reaches the time limit
of 144 hours, so no observations can be made for longer budgets. In smaller
instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.
%Overall, the improvement from the first  to the last
%solutions is 2-3 times larger in UMM than CEGO. %
%\MANUEL{But this is a bit misleading because UMM runs for 4 times more
%  evaluations, perhaps we should also measure up to the evaluation that CEGO
%  reaches so we report both values?}
%

@@

In this paper, we have introduced UMM, a population-based probabilistic
algorithm based on an unbalanced Mallows model. The algorithm is designed for
black-box combinatorial problems on permutation landscapes and when the budget
of fitness evaluations is severely limited (here, up to $400$ evaluations).
Our experiments presented here show that UMM is able to match the results of
CEGO, a state-of-the-art Bayesian optimizer for expensive black-box
optimization if the computation time of the optimizers is not a concern.

In addition, UMM is specially well-suited for budget-limited scenarios that are
still time-sensitive, e.g., when the overhead incurred by the optimizer should
be no more than a few minutes added to each fitness evaluation. By comparison,
CEGO may require up to one hour per fitness evaluation, with the time
increasing non-linearly with permutation size and larger budgets.  Hence, UMM
is a computationally feasible alternative for relatively large problem sizes.

%\MANUEL{UMM is able to model the LOP landscape
%  at least as well as CEGO as shown by the plots where UMM shows a linear
%  decrease in fitness.}  
%  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
%  solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q
%  es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos y
%  luego que addemás es más rápido} 
%  \MANUEL{Pero UMM es peor en rec13 y nug12, y
%  en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más
%  rápido. No podemos saber qué pasaría si CEGO llegara a 400. Podríamos medir
%  lo que pasa en el momento en que matamos a CEGO y dar también ese valor
%  entonces la diferencia en velocidad no importaría. } 
% The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
% CEGO  it sharply
% increases with $n$, requiring 
% sometimes more than hour per fitness evaluation as shown in the experimental section. 
% Hence, UMM is a computationally
% feasible alternative for relatively large problem sizes.

Although we were inspired by previous work on ACO for expensive black-box
combinatorial problems~\citep{PerLopStu2015si}, UMM is, to the best of our
knowledge, the first EDA specifically designed for such problems that is able
to match (and sometimes even improve) the results of a Bayesian
optimizer. Moreover, it is overwhelmingly faster than the alternatives for
large problem sizes.  The results presented here show that, despite the
intrinsic difficulty of the expensive black-box combinatorial scenario, there
are still significant advances yet to be made. Thus, we hope that this paper
will motivate further research. % on such problems.
%\MANUEL{I'm thinking about finishing here and commenting out the following paragraph}

An interesting result is that the behavior of UMM (and the same for CEGO) on
the PFSP (and some QAP instances) is close to random. The same behavior is
observed for CEGO, which seems to have been never reported until now.
This result may be due to the distance
metric used, but further analysis is needed to corroborate this
conjecture. Here, we have focused on Kendall's-$\tau$ distance, however, we are certain it is
possible to extend UMM to other distance metrics, which will allow us to
dynamically select among various distance metrics for an unknown black-box
% \ekhine{}{estoy totalmente en contra de esto :) }
permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
always obvious, thus a more detailed analysis would be needed to provide either
generally good static values or an online adaptation approach. Finally, we have
used here three very different combinatorial problems (LOP, PFSP and QAP) as
black-box benchmarks. However, an even more diverse range of problems would be
needed to understand the behavior of UMM on real-world black-box combinatorial
landscapes.

@@

%% MANUEL: Blind, so no acknowledgments yet.
\begin{smaller}
  \paragraph*{Reproducibility.} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
  \vspace*{-1em}
\paragraph*{Acknowledgements.}
% M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
%
%\manuel{}{: We should also thank the COST ACTION: Ask Carlos?} \ekhine{}{Sure!, esto no habra q ponerlo aun por ser blind, no?}
Thanks to: Hao Wang (Leiden University) for pointing us to the arguments of
\citep{EriPeaGar2019scalable}; Thomas Stützle (Université libre de Bruxelles) for suggesting works
about differences between QAPLIB instances; and Ahmed Boujaada for the experimental support. 
\end{smaller}

\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
\bibliographystyle{splncs04natnodoi}
\bibliography{optbib/abbrev,optbib/authors,optbib/abbrevshort,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}
